### 2020-11-09 manuscript/manuscript.Rmd
update

---
title: "BMI-manuscript"
output: 
    bookdown::pdf_document2:  
        keep_tex: true
bibliography: cite.bib
link_citations: true
---
\section{introduction}

Body Mass Index (BMI) plays an important rule in predicting heart disease risk[@katbody]. Our goal is to compare 90th quatile regrission for BMI in Diabtese and prediabtes population.

\section{Quantile regression}
Quantile regression is an important tool used to regress the dependent variable with high variance over the independent variables.The independent variabeles are Gender, Age, race, and statin use.

For a random variable $X$, the comulative distrubution function (CDF) is $$F(X)=P(X\leq x)$$, and the $\tau$th quantile of $X$ is defined by $$ F^{-1}(\tau)=\text{inf}\{x: F(x)\ge \tau\} $$ where $0<\tau<1$.
Let the loss function is defined by $$\rho_{\tau}(u)= u(\tau-I(u<0))$$  
[@koenker2005]. The expected loss function is defined by
\begin{equation}
E\rho_{\tau}(X-\hat{x})=
\end{equation}

Least square method solve $$ \underset{\mu}{\text{min}}\sum_{i=1}^n(y_i-\mu)^2$$

\section{methods}

prevalence and incidence as a function
of age (18 – 84 years in 1-year
intervals), race/ethnicity (non-Hispanic
white, non-Hispanic black, Hispanic, or
other), sex, and BMI (underweight,
 18.5 kg/m2; normal weight, 18.5 to
 25 kg/m2; overweight, 25 to 30 kg/
m2; obese, 30 to 35 kg/m2; and very
From the figure we see

#Data
 The data used in this study is National Health and Nutrition Examination Survey data (NHANES)[@NHANES]. The data has

```{r, out.width = "80%", fig.align = "center", fig.cap = "A better figure caption", echo = FALSE}
# Created by file exploratory-data-analysis.R
if (!file.exists(here::here("images", "exploratory.png"))) {
    source(here::here("exploratory-data-analysis.R"))
}
knitr::include_graphics(here::here("images", "exploratory.png"))
```
# References