### 2020-11-09 manuscript/manuscript.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={BMI-manuscript},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{BMI-manuscript}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{introduction}

Body Mass Index (BMI) plays an important rule in predicting heart
disease risk(Katzmarzyk et al. 2012). Our goal is to compare 90th
quatile regrission for BMI in Diabtese and prediabtes population.

(Van de Kassteele et al. 2012)

\section{methods}

prevalence and incidence as a function of age (18 -- 84 years in 1-year
intervals), race/ethnicity (non-Hispanic white, non-Hispanic black,
Hispanic, or other), sex, and BMI (underweight, 18.5 kg/m2; normal
weight, 18.5 to 25 kg/m2; overweight, 25 to 30 kg/ m2; obese, 30 to 35
kg/m2; and very From the figure we see

\section{Quantile regression}

Quantile regression is an important tool used to regress the dependent
variable with high variance over the independent variables.

For a random variable \(X\), the comulative distrubution function (CDF)
is \[F(X)=P(X\leq x)\], and the \(\tau\)th quantile of \(X\) is defined
by \[ F^{-1}(\tau)=\text{inf}\{x: F(x)\ge \tau\} \] where \(0<\tau<1\)

\section{Data}\label{data}

The data used in this study is National Health and Nutrition Examination
Survey data (NHANES)(Disease Control and (CDC) 2018)

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{C:/Users/Muhannad/Desktop/multvariat/R_codes/tonton/BMI/images/exploratory}

}

\caption{A better figure caption}\label{fig:unnamed-chunk-1}
\end{figure}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-NHANES}{}
Disease Control, Centers for, and Prevention (CDC). 2018. ``National
Health and Nutrition Examination Survey Data (Nhanes.''

\hypertarget{ref-katbody}{}
Katzmarzyk, Peter T, Bruce A Reeder, Susan Elliott, Michel R Joffres,
Punam Pahwa, Kim D Raine, Susan A Kirkland, and Gilles Paradis. 2012.
``Body Mass Index and Risk of Cardiovascular Disease, Cancer and
All-Cause Mortality.'' \emph{Canadian Journal of Public Health} 103 (2).
Springer: 147--51.

\hypertarget{ref-van2012estimating}{}
Van de Kassteele, Jan, RT Hoogenveen, PM Engelfriet, PHM Van Baal, and
HC Boshuizen. 2012. ``Estimating Net Transition Probabilities from
Cross-Sectional Data with Application to Risk Factors in Chronic Disease
Modeling.'' \emph{Statistics in Medicine} 31 (6). Wiley Online Library:
533--43.

\end{document}